---
title: "admin-page-conventions"
subtitle: "Standards derived from the `admin-page-conventions` skill"
---

## Purpose

Use this skill when working on admin (SDB/administrative records) pages in The Societal Mirror project. Triggers when user asks to add a chart, table, or section to pages like membership.qmd, centre_profile.qmd, or any page working with member_base_data, center data, enrollment data, or GCC constellation/region analysis from the Shambhala Database (SDB).

## Standard


# Admin Page Conventions

Admin pages draw from the Shambhala Database (SDB) — not survey responses. They use `member_base_data`, `base_center_data`, `center_member_count_class`, `membership_changes_year_detail`, and related datasets loaded by `shared_sdb_code_26.R`.

## Visual Identity: sdb_background_tint

Every admin plot MUST carry the SDB tint in both `plot.background` and `panel.background`. This visually distinguishes admin pages from survey pages.

```
theme(
  plot.background  = element_rect(fill = sdb_background_tint),  # "#F4F4F4"
  panel.background = element_rect(fill = sdb_background_tint)
)
```

- Survey (member) pages use `member_background_tint` (#FEF7F4)
- Survey (leader) pages use `leader_background_tint` (#F4FFF5)
- Admin/SDB pages use `sdb_background_tint` (#F4F4F4)

## mirror_theme(): Explicit vs Implicit

`shared_code_26.R` calls `theme_set(mirror_theme())`, so the theme is already the global default. Two valid styles appear across pages:

```
# Implicit (centre_profile.qmd style) — theme already set globally
ggplot(...) + geom_col() + theme(panel.grid.major.x = element_blank(), ...)

# Explicit (membership.qmd style) — redundant but harmless
ggplot(...) + geom_col() + mirror_theme() + theme(panel.grid.major.x = element_blank(), ...)
```

Both work. Be consistent within a page.

## Color Constants

| Constant | Hex | Use |
|---|---|---|
| `dark_bar_green` | #488460 | Primary bar/area fill (single series) |
| `sdb_background_tint` | #F4F4F4 | Plot/panel background |
| `size_palette` (4 colors) | greens | Fill when colored by center size group |
| `size_palette` (2 colors) | light + dark green | Fill for year-comparison (last yr / this yr) |
| `membership_colors[2:3]` | #86BD9C / #7994B8 | Member vs Friend two-series fills |

### size_palette — Two Variants

```
# 4-color (stacked by size group) — defined globally in shared_sdb_code_26.R
size_palette <- c("#B1CEBD", "#7CA98E", "#488460", "#346046")

# 2-color (year comparison) — redefine locally in the chunk
size_palette <- c("#B1CEBD", "#488460")   # light = last year, dark = current year
```

### Member/Friend Color Constants

```
membership_colors <- c("#F4D786", "#86BD9C", "#7994B8")
names(membership_colors) <- c("Non member", "Member", "Past member")

member_friends_colors <- membership_colors[2:3]
names(member_friends_colors) <- c("Members", "Friends")

scale_fill_manual(values = member_friends_colors)
```

Use `fct_recode()` to align factor levels with color names:
```
mutate(membertype = fct_recode(membertype,
  Members = "current_members",
  Friends = "current_friends"
))
```

## Key Pre-built Objects (from shared_sdb_code_26.R)

Do not recompute these — they are already loaded:

| Object | Contents |
|---|---|
| `center_member_count_class` | One row per center: `center`, `active_members`, `size_group`, `gcc_region`, `societal_mirror_label`, `iscenter`, `center_visible`, `location_type`, `Year` |
| `center_member_count_class_last_year` | Same structure, prior year |
| `center_size_member_count` | Centers by `societal_mirror_label` × `size_group` (current year, urban active only) |
| `center_size_member_count_last_year` | Same, prior year |
| `center_size_comparison` | `bind_rows` of last + current year with factor-ordered `societal_mirror_label` |
| `gcc_order` | Character vector of `societal_mirror_label` ordered by `gcc_region` — use with `fct_relevel()` |
| `enrollment_location` | `location`, `calendar_year`, `enrollment_at_location`, `enrollment_from_location`, `net_enrollment` |
| `statistiques` | Historical member counts (POSIXct `created`, integer `members`) |
| `change_tracking` | Derived from `membership_changes_year_detail`: member/friend gains and losses by center × year |

## Key Data Pipelines

### Members / Friends snapshot
```
member_base_data |>
  filter(membertype == "Member")           # or "Friend of Shambhala"
  # or filter(membertype != "Non-member")  # members + friends combined
```

### By generation
```
member_base_data |>
  filter(membertype == "Member") |>
  count(imputed_generation, name = "count") |>
  mutate(
    pct   = count / sum(count) * 100,
    label = paste0(round(pct, 0), "%"),
    imputed_generation = fct_rev(imputed_generation)
  )
```

### By center size
```
center_member_count_class |>
  filter_out(center %in% c(5, 68)) |>
  filter(str_detect(location_type, "Meditation"), center_visible == 1) |>
  count(size_group, wt = active_members, name = "active_members") |>
  filter(!is.na(size_group)) |>
  mutate(size_group = fct_rev(size_group))
```

### By GCC constellation / region
```
center_member_count_class |>
  filter(Year == filter_year, center_visible == 1) |>
  count(gcc_region, societal_mirror_label, wt = active_members, name = "active_members") |>
  mutate(societal_mirror_label = fct_reorder(societal_mirror_label, gcc_region)) |>
  filter(!is.na(societal_mirror_label))
# Axis: scale_x_discrete(labels = scales::label_wrap(30))
```

### Year-to-year comparison pipeline
```
# Build current + prior year separately, then bind
current_yr <- center_member_count_class |>
  filter_out(center %in% c(5, 68)) |>
  filter(str_detect(location_type, "Meditation"), center_visible == 1) |>
  count(size_group, name = "active_centers") |>    # or wt = active_members
  filter(!is.na(size_group)) |>
  mutate(Year = as.character(filter_year))

last_yr <- center_member_count_class_last_year |>
  filter_out(center %in% c(5, 68)) |>
  filter(str_detect(location_type, "Meditation"), center_visible == 1) |>
  count(size_group, name = "active_centers") |>
  mutate(Year = as.character(filter_year - 1))

combined <- bind_rows(last_yr, current_yr) |>    # last year first = legend order
  filter(!is.na(size_group))

# For GCC region (no urban filter):
bind_rows(last_yr, current_yr) |>
  mutate(
    societal_mirror_label = fct_relevel(societal_mirror_label, gcc_order),
    societal_mirror_label = fct_rev(societal_mirror_label)
  )
```

### Membership changes / gains-losses (since 2020)
```
change_tracking |>
  filter(change_year >= "2020-01-01") |>
  count(change_year, wt = new_members, name = "new_members")
```

### Enrollment activity (offering vs participating)
```
cutoff <- 3

enrollment_location |>
  filter(calendar_year %in% c(2020, 2025)) |>
  select(-net_enrollment) |>
  pivot_longer(
    cols = c(enrollment_at_location, enrollment_from_location),
    names_to = "where", values_to = "enrollment"
  ) |>
  mutate(where = case_when(
    where == "enrollment_at_location"   ~ "Offering",
    where == "enrollment_from_location" ~ "Participating"
  )) |>
  left_join(center_member_count_class, by = c("location" = "center")) |>
  filter(enrollment > cutoff, !is.na(size_group)) |>
  summarise(centers = n_distinct(location),
            .by = c(where, size_group, calendar_year))
```

### Joins to center metadata
```
left_join(
  center_member_count_class |>
    filter_out(center %in% c(5, 68)) |>
    filter(str_detect(location_type, "Meditation"), center_visible == 1) |>
    select(center, size_group, gcc_region, societal_mirror_label),
  by = "center"
)
```

## Standard Exclusions

- **`filter_out(center %in% c(5, 68))`** — exclude Shambhala Online (5) and Shambhala Global (68)
- **`filter(center_visible == 1)`** — active centers only
- **`filter(str_detect(location_type, "Meditation"))`** — urban centres/groups only
- **`filter(change_year >= "2020-01-01")`** — "since 2020" analyses

## ggplot Templates

### Horizontal Bar (most common)
```
data |>
  mutate(category = fct_rev(category)) |>
  ggplot(aes(category, count)) +
  geom_col(fill = dark_bar_green, width = 0.7) +
  geom_text(aes(label = scales::comma(count)),
            position = position_stack(vjust = 0.5),
            color = "white", fontface = "bold") +
  coord_flip() +
  scale_y_continuous(labels = scales::comma, limits = c(0, <max>)) +
  labs(title = "...", subtitle = "...", x = NULL, y = NULL, caption = "...") +
  mirror_theme() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.background    = element_rect(fill = sdb_background_tint),
    panel.background   = element_rect(fill = sdb_background_tint)
  )
```

### Vertical Bar / Time Series Bar
```
data |>
  ggplot(aes(year_col, count)) +
  geom_col(fill = dark_bar_green) +
  geom_text(aes(label = count),
            position = position_stack(vjust = 0.5),
            color = "white", fontface = "bold") +
  labs(title = "...", x = NULL, y = NULL) +
  mirror_theme() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.background    = element_rect(fill = sdb_background_tint),
    panel.background   = element_rect(fill = sdb_background_tint)
  )
```

### Dodged Bar (year comparison, vertical)
```
dodge_width <- 1    # or 0.9 for narrower gaps

combined |>
  ggplot(aes(x = size_group, y = active_centers, fill = Year, group = Year)) +
  geom_col(position = position_dodge(width = dodge_width)) +
  geom_text(
    aes(y = active_centers,
        label = scales::number(active_centers, accuracy = 1, big.mark = ",")),
    position = position_dodge(width = dodge_width),
    vjust = -0.5                             # labels above bars
  ) +
  scale_fill_manual(values = size_palette) +
  scale_x_discrete(labels = scales::label_wrap(12)) +
  scale_y_continuous(limits = c(0, <max>)) +
  labs(title = "...", x = NULL, y = NULL, fill = "Year") +
  theme(
    legend.position    = "bottom",
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = sdb_background_tint),
    panel.background   = element_rect(fill = sdb_background_tint)
  )
```

### Dodged Bar with coord_flip (GCC region year comparison)
```
#| fig-height: 12

combined |>
  filter(!is.na(societal_mirror_label)) |>
  ggplot(aes(x = societal_mirror_label, y = active_centers, fill = Year, group = Year)) +
  geom_col(position = position_dodge(width = dodge_width)) +
  geom_text(
    aes(y = active_centers + 0.5, label = scales::number(active_centers)),
    position = position_dodge(width = dodge_width)
  ) +
  coord_flip() +
  scale_fill_manual(values = size_palette) +
  scale_x_discrete(labels = scales::label_wrap(30)) +
  labs(title = "...", x = NULL, y = NULL, fill = "Year") +
  theme(
    legend.position    = "bottom",
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = sdb_background_tint),
    panel.background   = element_rect(fill = sdb_background_tint)
  )
```

### Stacked Bar by Size within GCC Region
```
#| fig-height: 12

center_size_member_count |>
  ggplot(aes(x = societal_mirror_label, y = active_centers, fill = size_group)) +
  geom_col(position = "stack") +
  geom_text(aes(label = active_centers),
            position = position_stack(vjust = 0.5),
            color = "white") +
  coord_flip() +
  scale_fill_manual(values = size_palette) +       # 4-color palette
  scale_x_discrete(labels = scales::label_wrap(30)) +
  labs(title = "...", x = NULL, y = NULL, fill = NULL) +
  theme(
    legend.position   = "bottom",
    legend.box.just   = "left",
    legend.margin     = margin(0, 0, 0, 0),
    legend.box.margin = margin(0, 0, 0, -150),     # align legend with bars
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = sdb_background_tint),
    panel.background   = element_rect(fill = sdb_background_tint)
  )
```

### Faceted Stacked Bar
```
data |>
  ggplot(aes(x_var, y_var, fill = size_group)) +
  geom_col() +
  geom_text(aes(label = y_var),
            position = position_stack(vjust = 0.5),
            color = "white", fontface = "bold", size = 3.5) +
  geom_text(                                        # totals above bars
    data = data |> summarise(total = sum(y_var), .by = c(x_var, facet_var)),
    aes(y = total, label = total, fill = NULL),
    vjust = -0.5, fontface = "bold"
  ) +
  facet_wrap(~ facet_var) +
  scale_fill_manual(values = ...) +
  scale_y_continuous(limits = c(0, <max>)) +
  guides(fill = guide_legend(nrow = 2)) +
  theme(
    legend.position    = "top",
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = sdb_background_tint),
    panel.background   = element_rect(fill = sdb_background_tint),
    strip.background   = element_rect(fill = sdb_background_tint),  # required for facets
    strip.text         = element_text(hjust = 0.5)
  )
```

### Time Series Area Chart
```
data |>
  ggplot(aes(x = date_col, y = count)) +
  geom_area(fill = membership_colors[2]) +
  scale_y_continuous(
    labels = scales::label_comma(),
    breaks = seq(0, 12000, 2000),
    expansion(mult = c(0, .05)),
    limits = c(0, 12000)
  ) +
  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
  # or: scale_x_date(date_breaks = "1 year", date_labels = "%Y") for Date columns
  labs(y = NULL, x = NULL, title = "...") +
  mirror_theme() +
  theme(
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = sdb_background_tint),
    panel.background   = element_rect(fill = sdb_background_tint)
  )
```

To annotate specific data points (odd years, max):
```
highlights <- data |>
  filter(
    (month(date_col) == 1 & day(date_col) == 1 & year(date_col) %% 2 == 1) |
    date_col == max(date_col) | date_col == min(date_col)
  )
max_point <- data |> filter(count == max(count))

# Add to ggplot:
geom_point(data = highlights, size = 2, shape = 25, fill = "black",
           position = position_nudge(y = 200)) +
geom_text(data = highlights, aes(label = scales::comma(count)),
          vjust = -2, size = 3) +
geom_point(data = max_point, size = 2, shape = 25, fill = "red",
           position = position_nudge(y = 200)) +
geom_text(data = max_point,
          aes(label = paste0("Max: ", scales::comma(count), "\n", as.Date(date_col))),
          vjust = -0.8, size = 3, color = "red")
```

### Diverging Bar (gains and losses)
```
data |>
  ggplot(aes(x = year_col, y = count, fill = type)) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = 0, color = "black", linewidth = 1) +
  geom_text(aes(label = abs(count)),
            position = position_stack(vjust = 0.5),
            color = "white", fontface = "bold", size = 4) +
  scale_fill_manual(
    values = c("gains" = dark_bar_green, "losses" = "gray60"),
    labels = c("gains" = "Gains", "losses" = "Losses")
  ) +
  scale_y_continuous(
    breaks = seq(-800, 400, 100),
    labels = function(x) scales::comma(abs(x)),    # show absolute values
    limits = c(-800, 400)
  ) +
  labs(title = "...", x = NULL, y = NULL, fill = NULL) +
  mirror_theme() +
  theme(
    legend.position    = "top",
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank()
  )
```

## Grid Line Rules

- **Horizontal bars** (`coord_flip()`): blank `panel.grid.major.y` and `.minor.y`
- **Vertical bars / time series**: blank `panel.grid.major.x`
- **Faceted charts**: also add `strip.background = element_rect(fill = sdb_background_tint)`

## scale_y_continuous: expansion() for Headroom

```
scale_y_continuous(
  labels = scales::comma,
  breaks = seq(0, 12000, 2000),
  expansion(mult = c(0, .05)),    # 0 padding at bottom, 5% at top
  limits = c(0, 12000)
)
```

## gt Tables

### Basic template
```
data |>
  gt() |>
  cols_label(col1 = md("Line 1<br>Line 2"), col2 = "Label") |>
  fmt_number(columns = where(is.numeric), decimals = 0, use_seps = TRUE) |>
  fmt_percent(columns = pct_col, decimals = 0) |>
  grand_summary_rows(
    columns = where(is.numeric),
    fns = list("Total" = ~ sum(., na.rm = TRUE)),
    fmt = list(~ fmt_number(., decimals = 0, use_seps = TRUE))
  ) |>
  mrr_gt_theme()
```

### Grouped rows with group + grand summaries
```
data |>
  group_by(group_col) |>
  gt(groupname_col = "group_col") |>
  fmt_number(columns = c("col1", "col2"), decimals = 0, use_seps = TRUE) |>
  fmt_percent(columns = pct_col, decimals = 0) |>
  cols_label(...) |>
  cols_width(1 ~ px(300), 2:5 ~ px(100)) |>
  summary_rows(                               # subtotal per row group
    columns = c("col1", "col2"),
    fns = list("Total" = ~ sum(., na.rm = TRUE)),
    fmt = list(~ fmt_number(., decimals = 0, use_seps = TRUE))
  ) |>
  grand_summary_rows(                         # overall total
    columns = c("col1", "col2"),
    fns = list("Grand Total" = ~ sum(., na.rm = TRUE)),
    fmt = list(~ fmt_number(., decimals = 0, use_seps = TRUE))
  ) |>
  mrr_gt_theme() |>
  tab_options(stub_row_group.border.width = px(10)) |>
  as_raw_html()                               # prevents render issues for complex grouped tables
```

**`summary_rows()`** — one subtotal per row group. **`grand_summary_rows()`** — one total at table bottom. Both can appear together; call separately per column set when formatting differs.

### Spanner columns (year × metric)
```
data |>
  pivot_wider(values_from = metric, names_from = c(category, year)) |>
  gt() |>
  cols_label(
    Offering_2020 = "Offering", Participating_2020 = "Participating",
    Offering_2025 = "Offering", Participating_2025 = "Participating"
  ) |>
  tab_spanner(label = md("2020"), columns = c(Offering_2020, Participating_2020)) |>
  tab_spanner(label = md("2025"), columns = c(Offering_2025, Participating_2025)) |>
  # Nested spanners — apply widest first:
  tab_spanner(label = md("Year-end<br>Count"),         columns = 2:3) |>
  tab_spanner(label = md("Change from<br>prior year"), columns = 4:7) |>
  grand_summary_rows(
    columns = where(is.numeric),
    fns = list("Total" = ~ sum(., na.rm = TRUE)),
    fmt = list(~ fmt_number(., decimals = 0, use_seps = TRUE))
  ) |>
  mrr_gt_theme()
```

## Page Structure

Every admin section follows the Graph + Table panel-tabset pattern:

```
## Section Title

`  :::: {.panel-tabset}

### Graph
```{r}
#| label: fig-descriptive-name
# ggplot code
```

### Table
```{r}
#| label: tbl-descriptive-name
# gt() |> mrr_gt_theme() code
```

`  ::::
```

- `knitr::knit_exit()` in a chunk stops rendering — use to hide unfinished sections during development
- `{=html} <!-- ... -->` blocks for editorial comments hidden from output
- `#| eval: false` on exploratory or upload chunks not ready to render
- Section inventory in an `eval: false` chunk at top is a useful tracking pattern

## Chunk Options

```
#| label: fig-descriptive-name    # fig- prefix for plots, tbl- for tables
#| echo: false
#| message: false
#| warning: false
#| fig-height: 12                 # always use for GCC region charts (18+ rows)
```

All chunks suppress output by default via the setup chunk.

